Estimating real-time high street footfall from Wi-Fi probe requests

### 2018-01-30 abstract.tex
final draft for paul and james

	\section{Abstract}

@@

	\begin{figure*}
		\begin{center}
			\includegraphics[width=0.8\textwidth]{outputs/clustering_1.png}
			\caption{Clustering probe requests based on increasing sequence
			numbers present in them.}
		\end{center}
	\end{figure*}

@@

	Then we turn our attention to other fields which can provide us information
	on the device which generated the probe request and hence enabling us to 
	finger print them uniquely.
	From a preliminary investigation, we find that the data fields - SSID and length of
	the tags are very sparse and not useful in identifying the unique device while 
	OUIs, length of the packet and sequence numbers can be crucial in doing the same.
	A pair of probes with the same OUI shows the device manufacturer and same length of the packet signifies that 
	there is a high probability that it was generated by a similar devcie.

	We first split the anonymised probe requests by the OUI and length then run a 
	graph based partition algorithm on them based on the sequence numbers. This 
	algortithm links probes with increasing sequence numbers in increasing time within a
	specific distance. It also enforces the rule that a probe request has only one 
	incoming and outgoing links. It then returns the membership of the probe requests in
	distinct connected clusters in the resulting graph thus classifying probe requests
	with sequentially increasing numbers as ones generated from the same device as shown in Figure 1. Finally
	all the individual subsets are joined and cleaned for repeating probe requests with
	unique clusters. This filtering in-turn reduces the difference of the sensor count to the manual count to 
	-18\%.

	The important detail to notice is that we have clustered and estimated unique device
	without the knowledge of the MAC address of the devices hence delivering output 
	without compromising the users' privacy. The methodology can be applications in 
	numerous footfall counting projects in retail, urban planning, facilities management etc.

>>>>

@@

	\begin{figure*}
		\begin{center}
			\includegraphics
				[width=0.8\textwidth,trim=4 4 4 4,clip]
				{outputs/clustering_2.png}
			\caption{Clustering probe requests based on increasing sequence
			numbers present in them.}
			label{fig}
		\end{center}
	\end{figure*}

@@

	We then use the fields - OUI, lengths and sequence number,
	to tackle the noise from devices which anonymising the probe requests.
	OUI and length are used to split the dataset into
	groups of probe requests from similar
	devices and each subset is classified further
	based on a graph based clustering algorithm 
	where each cluster corresponds to an unique device.
	The algorithm works by creating a graph where the probe requests
	are the nodes and the links are created
	between them based on the following rules,

	\begin{enumerate}
		\item A link can go only forward in time.
		\item A link can exist between nodes with a 
			maximum time difference of $\alpha$ (time threshold)
		\item A link can go from low to high sequence numbers.
		\item A link can exist between nodes with a
			maximum sequence number difference of $\beta$ (sequence threshold)
		\item A node can have only one incoming link and outgoing link
			which is the shortest of all possible such links.
	\end{enumerate}

	The nodes are then classified based on the unique connected component they
	belong to. This classification is assigned as the unique identifier for the
	anonymised probe requests.
	\ref{fig} shows the clustering process for where the black dots show
	probe requests and red lines connecting them show the clusters
	corresponding to the same device generated them.
	We finally combine both normal and anonymised probe requests, 
	aggregate them based on their unique identifier and remove
	repeating probe requests which reduces the difference between
	the sensor counts and the manual counts to -18\%.

	It is important to note that the filtering process is done based on just
	the information present in the probe requests
	and their temporal distribution.
	This makes sure that though the mobile devices are uniquely identified
	there is no more personal data is generated linking the probe requests
	to the users of the mobile devices. This method essentially gives us a
	way to estimate the footfall in real-time without identifying or tracking the
	mobile devices themselves.

	This Wi-Fi based footfall counting methodology offers a large number of
	applications and benefits. Since Wi-Fi based sensors are inexpensive and the
	data model is scalable, it is readily possible to use this methodology for
	a large network of sensors to gather granular data on pedestrian footfall.
	Projects such as Smart Street Sensors, can utilise the methodology to
	overcome the challenges introduced by implementation of MAC address
	randomisation. Such precise and granular data also enables us in confidently
	model the pedestrian flow in urban road networks and will be indispensable
	tool in the smart city framework. This can also be used in understanding and
	classifying geographical areas based on the spatio temporal distribution of
	the volume of activity in them.
